Type Safe Redis Queries

### 2016-06-15 aplas16/sections/Embedding.tex
Up to assertions, and the final example.

A singleton type is a type that has only one term. When the term is built, it
carries a type that can be inspected by the type checker. The term can be think
of as a representative of its type at the realm of runtime values. For our
purpose, we will use the following type \ensuremath{\Conid{Proxy}}:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{data}\;\Conid{Proxy}\;\Varid{t}\mathrel{=}\Conid{Proxy}~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
For every type \ensuremath{\Varid{t}}, \ensuremath{\Conid{Proxy}\;\Varid{t}} is a type that has only one term: \ensuremath{\Conid{Proxy}}.
\footnote{While giving the same name to both the type and the term can be very
confusing, it is unfortunately a common practice in the Haskell community.}
To call \ensuremath{\Varid{del}}, instead of passing a key as a \ensuremath{\Conid{String}}, we give it a proxy with
a specified type:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{del}\;(\Conid{Proxy}\mathbin{::}\Conid{Proxy}\;\text{\tt \char34 A\char34})~~,{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
where \ensuremath{\text{\tt \char34 A\char34}} is not a value, but a string lifted to a type (of kind \ensuremath{\Conid{Symbol}}).
Now that the type checker has access to the key, the type of \ensuremath{\Varid{del}} could be
something alone the line of \ensuremath{\Varid{del}\mathbin{::}\Conid{Proxy}\;\Varid{s}\to \Conid{Edis}\;\Varid{xs}\;(\Conid{Del}\;\Varid{xs}\;\Varid{s})\mathbin{...}}.

@@

A final note: functions the \ensuremath{\Varid{encode}}, from the Haskell library {\sc cereal},
helps to convert certain datatype that are {\em serializable} into \ensuremath{\Conid{ByteString}}.
The function and its dual \ensuremath{\Varid{decode}} will be use more later.
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{encode}{}\<[9]%
\>[9]{}\mathbin{::}\Conid{Serialize}\;\Varid{a}\Rightarrow \Varid{a}\to \Conid{ByteString}~~,{}\<[E]%
\\
\>[B]{}\Varid{decode}{}\<[9]%
\>[9]{}\mathbin{::}\Conid{Serialize}\;\Varid{a}\Rightarrow \Conid{ByteString}\to \Conid{Either}\;\Conid{String}\;\Varid{a}~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

While the \ensuremath{\Varid{set}} command in \Hedis{} always writes a string to the data store,
the corresponding \ensuremath{\Varid{set}} in \Redis{} applies to any serializable type (those
in the class \ensuremath{\Conid{Serialize}}), and performs the encoding for the user:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{6}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{set}{}\<[6]%
\>[6]{}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Serialize}\;\Varid{x}){}\<[E]%
\\
\>[6]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Varid{x}\to \Conid{Edis}\;\Varid{xs}\;(\Conid{Set}\;\Varid{xs}\;\Varid{s}\;(\Conid{StringOf}\;\Varid{x}))\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Status}){}\<[E]%
\\
\>[B]{}\Varid{set}\;\Varid{key}\;\Varid{val}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.set}\;(\Varid{encodeKey}\;\Varid{key})\;(\Varid{encode}\;\Varid{val})~~,{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
where the function \ensuremath{\Varid{encode}\mathbin{::}\Conid{Serialize}\;\Varid{a}\Rightarrow \Varid{a}\to \Conid{ByteString}}

For example, executing \ensuremath{\Varid{set}\;(\Conid{Proxy}\mathbin{::}\Conid{Proxy}\;\text{\tt \char34 A\char34})\;\Conid{True}} updates the dictionary
with an entry \ensuremath{\mbox{\textquotesingle}(\text{\tt \char34 A\char34},\Conid{StringOf}\;\Conid{Bool})}: if \ensuremath{\text{\tt \char34 A\char34}} is not in the dictionary,
the entry is added; otherwise the old type of \ensuremath{\text{\tt \char34 A\char34}} is updated to
\ensuremath{\Conid{StringOf}\;\Conid{Bool}}.

\Redis{} command \texttt{INCR} reads the string associated with the given key,
parse it as an integer, and increments it, before storing it back. The command
\texttt{INCRBYFLOAT} increments the floating number, associated to the given
key. They can be
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{incr}{}\<[7]%
\>[7]{}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Get}\;\Varid{xs}\;\Varid{s}\mathord{\sim}\Conid{Just}\;(\Conid{StringOf}\;\Conid{Integer})){}\<[E]%
\\
\>[7]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Conid{Edis}\;\Varid{xs}\;\Varid{xs}\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Integer}){}\<[E]%
\\
\>[B]{}\Varid{incr}\;\Varid{key}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.incr}\;(\Varid{encodeKey}\;\Varid{key})~~,{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{incrbyfloat}{}\<[14]%
\>[14]{}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Get}\;\Varid{xs}\;\Varid{s}\mathord{\sim}\Conid{Just}\;(\Conid{StringOf}\;\Conid{Double})){}\<[E]%
\\
\>[14]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Conid{Double}\to \Conid{Edis}\;\Varid{xs}\;\Varid{xs}\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Double}){}\<[E]%
\\
\>[B]{}\Varid{incrbyfloat}\;\Varid{key}\;\Varid{eps}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.incrbyfloat}\;(\Varid{encodeKey}\;\Varid{key})\;\Varid{eps}~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
Notice the use of (\ensuremath{\mathord{\sim}}), \emph{equality constraints}~\cite{typeeq}, to enforce
that the intended type of value associated with key \ensuremath{\Varid{s}} must respectively be
\ensuremath{\Conid{Integer}} and \ensuremath{\Conid{Double}}.

@@

In the following example, \texttt{LPUSH} appends a string \texttt{"bar"} to the
list associated to \texttt{some-list}, or creates a list if \texttt{some-list}
does not yet present. The command \texttt{LLEN} then returns the length of
the list. If the key passed to \texttt{LLEN} is not associated to a list,
it signals an error.
\begin{tabbing}\tt
~redis\char62{}~LPUSH~some\char45{}list~bar\\
\tt ~\char40{}integer\char41{}~1\\
\tt ~redis\char62{}~LLEN~some\char45{}list\\
\tt ~\char40{}integer\char41{}~1\\
\tt ~redis\char62{}~SET~some\char45{}string~foo\\
\tt ~OK\\
\tt ~redis\char62{}~LLEN~some\char45{}string\\
\tt ~\char40{}error\char41{}~WRONGTYPE~Operation~against~a~key~holding\\
\tt ~the~wrong~kind~of~value
\end{tabbing}
One may give \ensuremath{\Varid{llen}} this type:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{llen}{}\<[7]%
\>[7]{}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Get}\;\Varid{xs}\;\Varid{s}\mathord{\sim}\Conid{Just}\;(\Conid{ListOf}\;\Varid{x})){}\<[E]%
\\
\>[7]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Conid{Edis}\;\Varid{xs}\;\Varid{xs}\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Integer}){}\<[E]%
\\
\>[B]{}\Varid{llen}\;\Varid{key}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.llen}\;(\Varid{encodeKey}\;\Varid{key})~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
However, this is not an accurate specification of \texttt{LLEN} in \Redis{} ---
\texttt{LLEN} also accepts keys that do not exist, and replies with \texttt{0}:
\begin{tabbing}\tt
~redis\char62{}~LLEN~nonexistent\\
\tt ~\char40{}integer\char41{}~0
\end{tabbing}
What we wish to have in the class constraint of \ensuremath{\Varid{llen}} is thus a predicate
equivalent to \ensuremath{\Conid{Get}\;\Varid{xs}\;\Varid{s}\doubleequals\Conid{Just}\;(\Conid{ListOf}\;\Varid{x})\mathrel{\vee}\neg \;(\Conid{Member}\;\Varid{xs}\;\Varid{s})}. The situation is the same with \ensuremath{\Varid{lpush}}: it returns normally either when
the key is associated to a list or does not exist at all, while signaling an
error when the associated value is not a list.

Unfortunately, expressing disjunctions in constraints is much more difficult
than expressing conjunctions. To impose a class constraint \ensuremath{\Conid{P}\mathrel{\wedge}\Conid{Q}}, one may
simply put them both in the type: \ensuremath{(\Conid{P},\Conid{Q})\Rightarrow \mathbin{...}}. To express top-level
constraints

@@

We could achieve this simply by translating the semantics we want to the
 domain of Boolean, with type-level boolean functions such as
\ensuremath{(\mathrel{\wedge})},
\ensuremath{(\mid )}, \ensuremath{\Conid{Not}},
\ensuremath{(\doubleequals)}, etc.\footnotemark To avoid

\footnotetext{Available in \text{Data.Type.Bool} and
 \text{Data.Type.Equality}}

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{Get}\;\Varid{xs}\;\Varid{s}\doubleequals\Conid{Just}\;(\Conid{ListOf}\;\Varid{x})\mathrel{\vee}\Conid{Not}\;(\Conid{Member}\;\Varid{xs}\;\Varid{s}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

To avoid addressing the type of value (as it may not exist at all), we defined
 an auxiliary predicate \ensuremath{\Conid{IsList}\mathbin{::}\Conid{Maybe}\mathbin{*}\to \Conid{Bool}} to
 replace the former part.

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{IsList}\;(\Conid{Get}\;\Varid{xs}\;\Varid{s})\mathrel{\vee}\Conid{Not}\;(\Conid{Member}\;\Varid{xs}\;\Varid{s}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

The type expression above has kind \ensuremath{\Conid{Bool}}, we could make it
 a type constraint by asserting equality.

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}(\Conid{IsList}\;(\Conid{Get}\;\Varid{xs}\;\Varid{s})\mathrel{\vee}\Conid{Not}\;(\Conid{Member}\;\Varid{xs}\;\Varid{s}))\mathord{\sim}\Conid{True}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

With \emph{constraint kind}, a recent addition to GHC, type constraints now has
 its own kind: \ensuremath{\Conid{Constraint}}. That means type constraints
 are not restricted to the left side of a \ensuremath{\Rightarrow } anymore,
 they could appear in anywhere that accepts something of kind
 \ensuremath{\Conid{Constraint}}, and any type that has kind
 \ensuremath{\Conid{Constraint}} can also be used as a type constraint.
 \footnote{See \url{https://downloads.haskell.org/~ghc/7.4.1/docs/html/users_guide/constraint-kind.html}.}

As many other list-related commands also have this ``List or nothing'' semantics,
 we could abstract the lengthy type constraint above and give it an alias with
 type synonym.

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{ListOrNX}\;\Varid{xs}\;\Varid{s}\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{5}{}\<[5]%
\>[5]{}(\Conid{IsList}\;(\Conid{Get}\;\Varid{xs}\;\Varid{s})\mathrel{\vee}\Conid{Not}\;(\Conid{Member}\;\Varid{xs}\;\Varid{s}))\mathord{\sim}\Conid{True}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

The complete implementation of \text{LLEN} with
\ensuremath{\Conid{ListOrNX}} would become:

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{llen}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{ListOrNX}\;\Varid{xs}\;\Varid{s}){}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Rightarrow \Conid{Proxy}\;\Varid{s}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\to \Conid{Edis}\;\Varid{xs}\;\Varid{xs}\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Integer}){}\<[E]%
\\
\>[B]{}\Varid{llen}\;\Varid{key}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.llen}\;(\Varid{encodeKey}\;\Varid{key}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

>>>>

A singleton type is a type that has only one term. When the term is built, it
carries a type that can be inspected by the type checker. The term can be think
of as a representative of its type at the realm of runtime values. For our
purpose, we will use the following type \ensuremath{\Conid{Proxy}}:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{data}\;\Conid{Proxy}\;\Varid{t}\mathrel{=}\Conid{Proxy}~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
For every type \ensuremath{\Varid{t}}, \ensuremath{\Conid{Proxy}\;\Varid{t}} is a type that has only one term: \ensuremath{\Conid{Proxy}}.%
\footnote{While giving the same name to both the type and the term can be very
confusing, it is unfortunately a common practice in the Haskell community.}
To call \ensuremath{\Varid{del}}, instead of passing a key as a \ensuremath{\Conid{String}}, we give it a proxy with
a specified type:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{del}\;(\Conid{Proxy}\mathbin{::}\Conid{Proxy}\;\text{\tt \char34 A\char34})~~,{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
where \ensuremath{\text{\tt \char34 A\char34}} is not a value, but a string lifted to a type (of kind \ensuremath{\Conid{Symbol}}).
Now that the type checker has access to the key, the type of \ensuremath{\Varid{del}} could be
something alone the line of \ensuremath{\Varid{del}\mathbin{::}\Conid{Proxy}\;\Varid{s}\to \Conid{Edis}\;\Varid{xs}\;(\Conid{Del}\;\Varid{xs}\;\Varid{s})\mathbin{...}}.

@@

A final note: the function \ensuremath{\Varid{encode}}, from the Haskell library {\sc cereal},
helps to convert certain datatype that are {\em serializable} into \ensuremath{\Conid{ByteString}}.
The function and its dual \ensuremath{\Varid{decode}} will be use more later.
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{encode}{}\<[9]%
\>[9]{}\mathbin{::}\Conid{Serialize}\;\Varid{a}\Rightarrow \Varid{a}\to \Conid{ByteString}~~,{}\<[E]%
\\
\>[B]{}\Varid{decode}{}\<[9]%
\>[9]{}\mathbin{::}\Conid{Serialize}\;\Varid{a}\Rightarrow \Conid{ByteString}\to \Conid{Either}\;\Conid{String}\;\Varid{a}~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

While the \ensuremath{\Varid{set}} command in \Hedis{} always writes a string to the data store,
the corresponding \ensuremath{\Varid{set}} in \Redis{} applies to any serializable type (those
in the class \ensuremath{\Conid{Serialize}}), and performs the encoding for the user:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{6}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{set}{}\<[6]%
\>[6]{}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Serialize}\;\Varid{x}){}\<[E]%
\\
\>[6]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Varid{x}\to \Conid{Edis}\;\Varid{xs}\;(\Conid{Set}\;\Varid{xs}\;\Varid{s}\;(\Conid{StringOf}\;\Varid{x}))\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Status}){}\<[E]%
\\
\>[B]{}\Varid{set}\;\Varid{key}\;\Varid{val}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.set}\;(\Varid{encodeKey}\;\Varid{key})\;(\Varid{encode}\;\Varid{val})~~,{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
For example, executing \ensuremath{\Varid{set}\;(\Conid{Proxy}\mathbin{::}\Conid{Proxy}\;\text{\tt \char34 A\char34})\;\Conid{True}} updates the dictionary
with an entry \ensuremath{\mbox{\textquotesingle}(\text{\tt \char34 A\char34},\Conid{StringOf}\;\Conid{Bool})}. If \ensuremath{\text{\tt \char34 A\char34}} is not in the dictionary,
this entry is added; otherwise the old type of \ensuremath{\text{\tt \char34 A\char34}} is updated to
\ensuremath{\Conid{StringOf}\;\Conid{Bool}}.

\Redis{} command \texttt{INCR} reads the string associated with the given key,
parse it as an integer, and increments it, before storing it back. The command
\texttt{INCRBYFLOAT} increments the floating number, associated to the given
key. They can be
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{incr}{}\<[7]%
\>[7]{}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Get}\;\Varid{xs}\;\Varid{s}\mathord{\sim}\Conid{Just}\;(\Conid{StringOf}\;\Conid{Integer})){}\<[E]%
\\
\>[7]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Conid{Edis}\;\Varid{xs}\;\Varid{xs}\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Integer}){}\<[E]%
\\
\>[B]{}\Varid{incr}\;\Varid{key}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.incr}\;(\Varid{encodeKey}\;\Varid{key})~~,{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{incrbyfloat}{}\<[14]%
\>[14]{}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Get}\;\Varid{xs}\;\Varid{s}\mathord{\sim}\Conid{Just}\;(\Conid{StringOf}\;\Conid{Double})){}\<[E]%
\\
\>[14]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Conid{Double}\to \Conid{Edis}\;\Varid{xs}\;\Varid{xs}\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Double}){}\<[E]%
\\
\>[B]{}\Varid{incrbyfloat}\;\Varid{key}\;\Varid{eps}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.incrbyfloat}\;(\Varid{encodeKey}\;\Varid{key})\;\Varid{eps}~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
Notice the use of (\ensuremath{\mathord{\sim}}), \emph{equality constraints}~\cite{typeeq}, to enforce
that the intended type of value associated with key \ensuremath{\Varid{s}} must respectively be
\ensuremath{\Conid{Integer}} and \ensuremath{\Conid{Double}}. The function \ensuremath{\Varid{incr}} is allowed to be called only
in a context where the type checker is able to reduce \ensuremath{\Conid{Get}\;\Varid{xs}\;\Varid{s}} to
\ensuremath{\Conid{Just}\;(\Conid{StringOf}\;\Conid{Integer})}. Similarly with \ensuremath{\Varid{incrbyfloat}}.

@@

Recall, from Section \ref{sec:introduction}, that commands \texttt{LPUSH key
val} and \texttt{LLEN key} returns normally either when \ensuremath{\Varid{key}} does not present
in the data store, or when \ensuremath{\Varid{key}} presents and is associated to a list.
What we wish to have in their constraint is thus a predicate equivalent to \ensuremath{\Conid{Get}\;\Varid{xs}\;\Varid{s}\doubleequals\Conid{Just}\;(\Conid{ListOf}\;\Varid{x})\mathrel{\vee}\neg \;(\Conid{Member}\;\Varid{xs}\;\Varid{s})}.

To impose a conjunctive constraint \ensuremath{\Conid{P}\mathrel{\wedge}\Conid{Q}}, one may simply put them both in the
type: \ensuremath{(\Conid{P},\Conid{Q})\Rightarrow \mathbin{...}}. Expressing disjunctive constraints is only slightly
harder, thanks to our type-level functions. Various operators for type-level
boolean and equality are defined in \text{Data.Type.Bool} and
\text{Data.Type.Equality}, like how we defined \ensuremath{\Conid{Or}} in Section
\ref{sec:type-fun}. We may thus write the predicate as:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{Get}\;\Varid{xs}\;\Varid{s}\doubleequals\Conid{Just}\;(\Conid{ListOf}\;\Varid{x})\mathbin{`\Conid{Or}`}\Conid{Not}\;(\Conid{Member}\;\Varid{xs}\;\Varid{s})~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
To avoid referring to \ensuremath{\Varid{x}}, which might not exist, we define an auxiliary predicate \ensuremath{\Conid{IsList}\mathbin{::}\Conid{Maybe}\mathbin{*}\to \Conid{Bool}}:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{type}\;\Varid{family}\;\Conid{IsList}\;(\Varid{x}\mathbin{::}\mathbin{*})\mathbin{::}\Conid{Bool}\;\mathbf{where}{}\<[E]%
\\
\>[B]{}\hsindent{5}{}\<[5]%
\>[5]{}\Conid{IsList}\;(\Conid{ListOf}\;\Varid{n})\mathrel{=}\mbox{\textquotesingle}\Conid{True}{}\<[E]%
\\
\>[B]{}\hsindent{5}{}\<[5]%
\>[5]{}\Conid{IsList}\;\Varid{x}{}\<[23]%
\>[23]{}\mathrel{=}\mbox{\textquotesingle}\Conid{False}~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
As many other list-related commands also have this ``List or nothing'' semantics, we give the type constraint a name:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{ListOrNX}\;\Varid{xs}\;\Varid{s}\mathrel{=}(\Conid{IsList}\;(\Conid{Get}\;\Varid{xs}\;\Varid{s})\mathrel{\vee}\Conid{Not}\;(\Conid{Member}\;\Varid{xs}\;\Varid{s}))\mathord{\sim}\Conid{True}~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\noindent The complete implementation of \text{LLEN} with
\ensuremath{\Conid{ListOrNX}} is therefore:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{lpush}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Serialize}\;\Varid{x},\Conid{ListOrNX}\;\Varid{xs}\;\Varid{s}){}\<[E]%
\\
\>[B]{}\hsindent{7}{}\<[7]%
\>[7]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Varid{x}\to \Conid{Edis}\;\Varid{xs}\;(\Conid{Set}\;\Varid{xs}\;\Varid{s}\;(\Conid{ListOf}\;\Varid{x}))\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Integer}){}\<[E]%
\\
\>[B]{}\Varid{lpush}\;\Varid{key}\;\Varid{val}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Redis}.lpush}\;(\Varid{encodeKey}\;\Varid{key})\;[\mskip1.5mu \Varid{encode}\;\Varid{val}\mskip1.5mu]~~,{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{llen}{}\<[7]%
\>[7]{}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{ListOrNX}\;\Varid{xs}\;\Varid{s}){}\<[E]%
\\
\>[7]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Conid{Edis}\;\Varid{xs}\;\Varid{xs}\;(\Conid{Either}\;\Conid{Reply}\;\Conid{Integer}){}\<[E]%
\\
\>[B]{}\Varid{llen}\;\Varid{key}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{\Conid{Hedis}.llen}\;(\Varid{encodeKey}\;\Varid{key})~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

% The type expression above has kind |Bool|, we could make it
%  a type constraint by asserting equality.
%
% With \emph{constraint kind}, a recent addition to GHC, type constraints now has
%  its own kind: |Constraint|. That means type constraints
%  are not restricted to the left side of a |=>| anymore,
%  they could appear in anywhere that accepts something of kind
%  |Constraint|, and any type that has kind
%  |Constraint| can also be used as a type constraint.
%  \footnote{See \url{https://downloads.haskell.org/~ghc/7.4.1/docs/html/users_guide/constraint-kind.html}.}

\subsection{Assertions}
\label{sec:assertions}

Finally, the creation/update behavior of \Redis{} functions is, in our opinion,
very error-prone. It might be preferable if we can explicit declare some new
keys, after ensure that they do not already exist (in our types), and renounce
them and forbid further access when we are sure that they shall not be referred
to anymore. This can be done below:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{declare}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Member}\;\Varid{xs}\;\Varid{s}\mathord{\sim}\Conid{False}){}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Conid{Proxy}\;\Varid{x}\to \Conid{Edis}\;\Varid{xs}\;(\Conid{Set}\;\Varid{xs}\;\Varid{s}\;\Varid{x})\;(){}\<[E]%
\\
\>[B]{}\Varid{declare}\;\Varid{s}\;\Varid{x}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{return}\;()~~,{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{renounce}\mathbin{::}(\Conid{KnownSymbol}\;\Varid{s},\Conid{Member}\;\Varid{xs}\;\Varid{s}\mathord{\sim}\Conid{True}){}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Rightarrow \Conid{Proxy}\;\Varid{s}\to \Conid{Edis}\;\Varid{xs}\;(\Conid{Del}\;\Varid{xs}\;\Varid{s})\;(){}\<[E]%
\\
\>[B]{}\Varid{renounce}\;\Varid{s}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{return}\;()~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
The command \ensuremath{\Varid{declare}\;\Varid{s}\;\Varid{x}} adds a new key \ensuremath{\Varid{s}} with type \ensuremath{\Varid{x}} to the dictionary,
if it does not already exist. Dually, \ensuremath{\Varid{renounce}} removes a key from the
dictionary. Even though it may still exist in the data store, it is not allowed
to be referred to. The command \ensuremath{\Varid{start}} initializes the dictionary to \ensuremath{\mbox{\textquotesingle}[~]}:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{start}\mathbin{::}\Conid{Edis}\;\mbox{\textquotesingle}[~]\;\mbox{\textquotesingle}[~]\;(){}\<[E]%
\\
\>[B]{}\Varid{start}\mathrel{=}\Conid{Edis}\mathbin{\$}\Varid{return}\;()~~.{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

\subsection{A Larger Example}
